Storm on Multi-core

### 2015-03-02 sapm/response.md

# RE: Distributed Software Development, and why I don't like it.

I enjoyed reading your article and given your experience I can see why you don't like DSD. However, in this response I would like to touch on some things that I believe might not necessarily have been the fault of DSD.

Firstly, your comments on team spirit resonated with me and as someone who is fairly open to DSD I agree that this is one of its disadvantages. I believe, however, that if your "local" team is large enough or you work in an office-share this can be helped. Additionally, I think it is fair to say that remote working is not for everyone.

Secondly, there is the language barrier problem. I think this is a really big lapse in judgement by the management of the company. I understand that the team was spread across the globe but if your management cannot ensure that you all develop in the same language then maybe they should have enforced regular code reviews (which I would advocate for anyway).

The reason why I claim that is because I was part of a  DSD team and we didn't have that problem at all. I worked at a company where some of the work for our main library was outsourced to a consultancy in Russia (the company could not scale quickly enough). The management set clear coding standards and the developers from Russia had to develop code in English (including comments). Moreover, they went through a testing period where my team would review their code and sign off on it.

Thirdly, I think that SLOC is a terrible metric. Period. It promotes bad practices such as copy-pasting code, avoiding re-factoring, and disadvantages people who write clean code. Additionally, developers could start optimising for SLOC and that could lead to unneeded complexity [ref]. Instead of measuring SLOC why not use some bug tracking or project management software e.g. Trello, Fogbugz, or Jira.

I believe the way they tracked work at the company I worked for was by daily stand-up meetings and going trough comments on cases that we were supposed to update regularly. I would imagine this takes up more time than counting SLOC but gives you a better idea of how much work people did.

> Was it fair? Was the developer really intentionally slacking? Who knows.

I feel that since you say "who knows" it is a little bit unfair to use it as part of your argument. There could have been valid reasons for that person being fired and I think it's a little bit unfair to suggest otherwise unless you know for sure.

Finally, I think that having your (project) manager start working 8 hours after you is less then ideal. There are limits to how distributed your team should be and I think there should definitely be a larger time overlap between you and your supervisor. Otherwise the communication will become the bottleneck.

http://en.wikipedia.org/wiki/Source_lines_of_code#Disadvantages

### 2015-03-02 proj/report.tex

%\pagenumbering{arabic}

@@

The document structure should include:
\begin{itemize}
\item
The title page  in the format used above.
\item
An optional acknowledgements page.
\item
The table of contents.
\item
The report text divided into chapters as appropriate.
\item
The bibliography.
\end{itemize}

Commands for generating the title page appear in the skeleton file and
are self explanatory.
The file also includes commands to choose your report type (project
report, thesis or dissertation) and degree.
These will be placed in the appropriate place in the title page.

The default behaviour of the documentclass is to produce documents typeset in
12 point.  Regardless of the formatting system you use,
it is recommended that you submit your thesis printed (or copied)
double sided.

The report should be printed single-spaced.
It should be 30 to 60 pages long, and preferably no shorter than 20 pages.
Appendices are in addition to this and you should place detail
here which may be too much or not strictly necessary when reading the relevant section.

\section{Using Sections}

Divide your chapters into sub-parts as appropriate.

\section{Citations}

Note that citations
(like \cite{P1} or \cite{P2})
can be generated using {\tt BibTeX} or by using the
{\tt thebibliography} environment. This makes sure that the
table of contents includes an entry for the bibliography.
Of course you may use any other method as well.

\section{Options}

There are various documentclass options, see the documentation.  Here we are
using an option ({\tt bsc} or {\tt minf}) to choose the degree type, plus:
\begin{itemize}
\item {\tt frontabs} (recommended) to put the abstract on the front page;
\item {\tt twoside} (recommended) to format for two-sided printing, with
  each chapter starting on a right-hand page;
\item {\tt singlespacing} (required) for single-spaced formating; and
\item {\tt parskip} (a matter of taste) which alters the paragraph formatting so that
paragraphs are separated by a vertical space, and there is no
indentation at the start of each paragraph.
\end{itemize}

\chapter{The Real Thing}

>>>>

\pagenumbering{arabic}

@@

In recent years, there has been an explosion of cloud computing software. After Google published their paper on MapReduce \cite{Anonymous:Jj3E6x7v}, many new open-source frameworks for distributed computation have emerged, most notably Apache Hadoop for batch processing and Apache Storm for real-time data stream processing.

The main idea of these frameworks is to split the work that needs to be carried out and distribute it across multiple nodes of a cluster. Commercial companies and researchers have been able to utilise these frameworks and create distributed systems \cite{5billion-sessions} which can accomplish things that would not be possible on a single computer. This has mostly been allowed by the low price of commodity hardware and good horizontal scaling properties.

% TODO: Here talk about What this paper is about !
% TODO: Maybe a section - move main idea here?
This project is about taking the ideas from the distributed system Apache Storm and applying them in the context of multi-core machines instead of clusters.

%% TODO: After that comes Motivation which explains why it's a good idea!
\section{Motivation}

While the cost of a commodity hardware cluster might be lower than the price of a single computer with equal power there are certain limitations:

\begin{itemize}

\item The nodes of a cluster communicate through network. This limits the speed of communication between processes that live on different nodes.

\item Distributed systems waste resources by replicating data to ensure reliability.

% TODO: Maybe mention renting vs owning here
\item Running a distributed computation on commodity hardware usually requires a data centre or renting out instances on cloud computing services such as Amazon EC2 or Rackspace. This is not ideal for some use cases which require full control over the system or a heightened level of security.

\end{itemize}

On the other hand, even though Moore’s law still holds true, processor makers now favour increasing the number of cores in CPU chips to increasing their frequency. This trend implies that the “free lunch” of getting better software performance by upgrading the processor is over and programmers now have to design systems with parallel architectures in mind. However, there are some limitations to this as well:

\begin{itemize}

\item It is generally believed that writing parallel software is hard. The traditional techniques of message passing and parallel threads sharing memory require the programmer to manage the concurrency at a fairly low level, either by using messages or locks.

%% TODO: Rephrase "nice if they could"
\item Apache Storm has become the de facto tool used in stream processing on a cluster and according to their "Powered By" page \cite{Anonymous:eikzOt4-} there are tens of companies already using Storm to process their real-time streams. It would be nice if they could keep that code.

\end{itemize}

\section{Main Idea}

The solution proposed in this paper is to take the existing Apache Storm project and port it for multi-core machines. This is implemented in a library Storm-MC with an API compatible with Apache Storm. This allows us to take an existing application written with Apache Storm in mind and run it in a multi-core setting. This way, we can avoid network latency and enjoy the significant performance improvements of a shared-memory environment.

\begin{itemize}

	\item Prices of high-end server have decreased and one can get a 32-core machine for 10,000 USD.

\end{itemize}

\section{Structure of the Report}
%% And after that explain the structure of the report :)

In chapter 1, blah blah.

\chapter{Background}

In this chapter we give background information necessary to understand the design of Storm-MC.

\section{Core Concepts}

There are several core concepts used by Storm and hence by extension Storm-MC as well. The concepts are put together to form a simple API that allows the user to break down a computation into separate components and then define how these components interact with each other.

\begin{description}
  \item[Topology] \hfill \\
  The programmer breaks down the computation to a network of components (called topology) which describes how the components interact with each other. Topology can always be represented as a directed acyclic graph. The topology is then submitted to the Storm library.
  \item[Spout] \hfill \\
  A spout is a source of a data-stream of a network. Typically, a spout reads from a message broker such as RabbitMQ or Kafka but can also generate its own stream or read from somewhere like the Twitter streaming API.
  \item[Bolt] \hfill \\
  A bolt takes a data stream as an input and transforms it into a different stream as its output. A bolt can perform a range of functions e.g. filter out tuples based on some criteria or perform a join of two different input streams.
\end{description}

\section{Storm Architecture}

The Storm cluster uses a Master-Worker pattern.

\subsection{Nimbus Node}

The master node (also called Nimbus) is responsible for the following tasks:

\begin{enumerate}
	\item Receive a submission with a topology and configuration from programmer.
	\item Distribute the work performed by different components to worker nodes.
\end{enumerate}

The Nimbus node is fault tolerant
%% TODO: Describe what Nimbus node does.

\subsection{Worker Nodes}

The worker nodes receive an assignment from the master node, spawn number of processes as specified by the configuration and do the actual stream processing.
%% TODO: Describe what worker nodes do.

\subsection{Zookeeper Nodes}

The Zookeeper nodes can be thought of as the glue of the network. They ensure that messages are delivered.
%% TODO: Describe how zookeeper nodes work.

\section{Inside a Worker Node}

The innards of a Worker node are as follows:

\begin{description}
	\item[Supervisor] \hfill \\
	A supervisor is a deamon process, run under supervision, that launches worker processes. This daemon is what the programmer runs on a worker node and it is the topmost layer.
	\item[Worker] \hfill \\
	A worker is a process that is assigned a port and listens to tuple messages on a socket associated with the port. Whenever it receives a tuple, it puts it on a queue where it is picked up by one or more executors of the worker process.
	\item[Executor] \hfill \\
	An executor controls the parallelism within a worker process. Every executor runs in a separate thread. Their job is to pick up tuples from the receiving queue, perform the function of a component they represent, and put the transformed tuples on the transfer queue, ready to be sent along to the next component. An executor runs one (the usual case) or more tasks.
	\item[Task] \hfill \\
	A task performs the actual data processing. However, within an executor thread all the tasks are executed sequentially. The main reason for having tasks is that the number of tasks stays the same throughout the lifetime of a topology but the number of executors can change (via rebalancing). Thus if some worker nodes in the cluster go down, the topology can continue executing with the same number of tasks as before.
\end{description}

\chapter{Design}

\chapter{Implementation}

\chapter{Evaluation}

\chapter{Conclusion}
